---
title: "Homicides_Scandinavian"
author: "Lea Skriver Hansen"
date: "1/7/2022"
output: 
  html_document: 
    toc: true 
    toc_float: true
  
---
*DISCLAIMER* 
this is not my first script with this data set. The first script went wrong and turned grey so i could not do anything with it so i had to delete it all and start over and that you can see in this new script 

to be able to knit the final results i put # with the codes that did not work 

#Introduction
this script is my final results including my thoughts and some of the mistakes i made throughout the process

# Downloading the relevant packages for this project 
```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(readxl)
```
the readxl package was downloaded because of issues with csv files (that will be explained later in the script)

#downloading the data
```{r}
#data <- read_excel("Digitale_metoder_projekt/Homicide_Murder_Rate.xls")

#Homicide_Murder_Rates <- read_excel("data/Homicide_Murder_Rates.xls")

```

from previus classes i have learned that csv files does not work for me so i tried putting xls file in by code, but that also did not work so i had to put in in without the code by importing the data 

#put the data in to ggplot failed attempt 
```{r}
#head(Homicide_Murder_Rates)
```

```{r}
#head(Homicide_Murder_Rates)
```
to check if the data was correct

#first try
```{r}
#ggplot(data = Homicide_Murder_Rates) +
 #geom_line(mapping = aes(x= Year,
                         # y= Homicide_Murder_Rates$Per_100_k_DK, )) + 
  #labs(x = "Year",
       #y = "Number of Murder per 100k",
       #title = "Murder Rate in Nothern Europe 1990-2018")
```

this was my first try to put it into ggplot but it said error
```{r}
#rlang::last_error()
```

trying to find out what the error was, but i could not figure it out 


```{r}
#second try 
#New_Homicide_Murder_Rates <- rename(Homicide_Murder_Rates,murder_per_100k = 'Murder rate in Nothern Europe (per 100,009)')
```
after looking at some of the things we have learned in class so i tried some of it but it still did not work

this was a try to see if i insertet one of the columns would work and it did but my goal is to get all the columns into a graphs so i can seperate them 
```{r}
#ggplot(data = Homicide_Murder_Rates) +
  #geom_line(mapping = aes(x= Year,
                          #y= Per_100_k_DK, )) + 
 # labs(x = "Year",
      # y = "Number of Murder per 100k",
      # title = "Murder Rate in Nothern Europe 1990-2018")
```
```{r}
#ggplot(data = Homicide_Murder_Rates) + 
  #geom_line(mapping = aes(x = Year, 
                          # y = Per_100_k_DK, 
                          # color = Entity)) +
  #facet_wrap(~ Entity, nrow = 2) +
  #labs(x = "Year",
       #y = "Number of Murder per 100k ",
       #title = "Murder Rate in Nothern Europe 1990-2018")
```
trying to see if i could separate them but of course that did not work when i only had one country


```{r}
#Homicide_Murder_Rates <- rename(Homicide_Murder_Rates, number_of_Murders_per_100k=n)
```
again trying anything to see if it would work but it did not 

#importing the new spreadsheet 
then i realized it was the spreadsheet that did not work and not just the code 

```{r}
#Murder_rates_Notherneurope_1990-2018 <- read_csv("Murder_rates_northeurope_1990-2018.csv")

#Murder_rates_Notherneurope_1990_2018 <- read_csv("Murder_rates_Notherneurope_1990-2018.csv")

#getwd()

#Murder_rates_Notherneurope_1990_2018 <- read_csv("Murder_rates_Notherneurope_1990-2018.csv")
```
again the code gave me trouble so i put the csv file without the code and by importing it 

#Checking the data is good to work with
```{r}
#head(Murder_rates_Notherneurope_1990_2018)
```

#prep the data for ggplot 
```{r}
#ggplot(data = Murder_rates_Notherneurope_1990_2018) +
 # geom_line(mapping = aes(x= Year,
                         # y= Murder_per_100k, )) + 
 # labs(x = "Year",
      # y = "Number of Murder per 100k",
      # title = "Murder Rate in Nothern Europe 1990-2018")
```

the data set in csv does not work, the data set is not separated probably so i have to use the xlsx file instead

```{r}
#Murder_rates_Notherneurope_1990_2018 <- read_excel("Homicides_Scandinavian/Murder_rates_Notherneurope_1990-2018.xlsx")

#getwd()

#Murder_rates_Notherneurope_1990_2018 <- read_excel("Murder_rates_Notherneurope_1990-2018.xlsx")
```

when trying to knit i had some issues because the data was importet and not by code so i tried using code
```{r}
Homicide_Murder_Rates <- read_excel("~/Desktop/Digitale metoder projekt /data/Homicide_Murder_Rates.xls")
```

the code could not show up because of the already existing one 
```{r}
rm(Murder_rates_Notherneurope_1990_2018)
```


after removing the importet data set i used code to insert the data and it was a success
```{r}
Murder_rates_Notherneurope_1990_2018 <- read_excel("Murder_rates_Notherneurope_1990-2018.xlsx")
#View(Murder_rates_Notherneurope_1990_2018)
```



# putting the data into ggplot
```{r}
ggplot(data = Murder_rates_Notherneurope_1990_2018) +
  geom_line(mapping = aes(x= Year,
                          y= Murder_per_100k, )) + 
  labs(x = "Year",
       y = "Number of Murder per 100k",
       title = "Murder Rate in Nothern Europe 1990-2018")
```

# seperate them by country ¨
```{r}
ggplot(data = Murder_rates_Notherneurope_1990_2018) + 
  geom_line(mapping = aes(x = Year, 
                           y = Murder_rates_Notherneurope_1990_2018$Murder_per_100k, 
                           color = Country)) +
  facet_wrap(~ Country, nrow = 3) +
  labs(x = "Year",
       y = "Murder per 100k",
       title = "Murder Rate in Nothern Europe 1990-2018")
```

#change layout 
```{r}
ggplot(data = Murder_rates_Notherneurope_1990_2018) + 
  geom_line(mapping = aes(x = Year, 
                           y = Murder_rates_Notherneurope_1990_2018$Murder_per_100k, 
                           color = Country)) +
  facet_wrap(~ Country, nrow = 2) +
  labs(x = "Year",
       y = "Murder per 100k",
       title = "Murder Rate in Nothern Europe 1990-2018")
```
#changing from geom_line to geom_bar
#first try
```{r}
#Murder_rates_Notherneurope_1990_2018
#ggplot( aes(x=Murder_per_100k)) + 
  #geom_bar()
```
this did not work because there was missing a chunk of code for it to work 
#second try
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Murder_per_100k)) +
  geom_bar()

```
realising i was missing something and when i added that to the code it worked 

#seperating them by country 
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Murder_per_100k)) +
  geom_bar(aes(fill= Country))
```
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Murder_per_100k)) +
  geom_bar(aes(fill= Country), position = "dodge")
```
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Murder_per_100k, y = Year, fill= Country)) + 
  geom_bar(stat = "identity", position = "dodge")
```
her i was a bit confused because it did not look rigth 
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Year, y = Murder_per_100k, fill= Country)) + 
  geom_bar(stat = "identity", position = "dodge")
```
then i realised that i had to switch the x and y and then it looked right 

#seperating them so each country have its own graph
```{r}
Murder_rates_Notherneurope_1990_2018 %>% 
  ggplot(aes(x = Year, y = Murder_per_100k, fill= Country)) + 
  geom_bar(stat = "identity", position = "dodge") +
  labs( title = "Murder rates 1990-2018", 
        x= "Year",
        y= "Murder per 100k populations") +
  facet_wrap(~ Country)
```

#find max in each country 
#Denmark 
```{r}
Per_100_k_DK <-c(1.01,1.24,0.98,1.09,1.31,0.85,0.84,0.82,0.76,0.85,0.98,0.71,0.53,0.98,0.8,1.21,0.89,0.97,1.09,0.98,0.92,1.67,1.29,1.13,1.44,1.37,1.2,1.32,0.8)

class(Per_100_k_DK)

max(Per_100_k_DK)
```

#Norway 
```{r}
Per_100_k_NO<-c(0.47,0.53,0.51,0.46,0.56,0.91,0.54,2.24,0.59,0.6,0.71,0.64,0.71,0.71,0.78,1.12,1.01,0.82,1.07,0.74,0.85,0.86,0.98,0.96,0.76,0.88,1.07,1.17,1.11)

class(Per_100_k_NO)

max(Per_100_k_NO)
```

#Sweden 
```{r}
Per_100_k_SW<-c(1.08,1.14,1.08,1.15,0.9,0.9,0.71,0.86,0.97,1,0.89,1.21,1,0.92,1.19,0.95,1.13,0.98,1.08,1.14,1.04,0.98,1.12,0.98,1.14,1.23,1.31,1.33,1.11)

class(Per_100_k_SW)

max(Per_100_k_SW)
```


```{r}
max(Per_100_k_SW)
```

#Finland
```{r}
Per_100_k_FI<-c(1.63,1.25,1.35,1.5,1.61,1.64,1.63,2.04,2.22,2.25,2.5,2.44,2.25,2.26,2.82,1.97,2.53,3,2.85,2.76,2.19,2.7,2.98,2.85,2.88,2.56,3.9,3.07,3.04)

class(Per_100_k_FI)

```
```{r}
max(Per_100_k_FI)
```

#Iceland 
```{r}
Per_100_k_IC<-c(0.89,0.9,0.3,0.91,0.61,0.31,0.31,0.93,0.62,0.32,0,1.02,1.03,0,1.4,0.35,1.78,0.72,0,0.73,0.37,0,0,0,0,0,0)

class(Per_100_k_IC)
```
```{r}
max(Per_100_k_IC)
```
```{r}
max(Per_100_k_IC)
```
```{r}
max(Per_100_k_IC)
```
i had a bit of trouble to find the right max with the data form Iceland but there was a mistake in c() so when i fixed that i got the right one

